### 2023-02-03 products/manuscript/Manuscript.qmd
Staging everything except data

---
title: "Evaluating Staffing Characteristics, Healthcare-associated Infections, and Provider Characteristics in U.S. Nursing Homes, 2021"
subtitle: ""
author: Kelly Hatfield
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

Part 1: Project Overview

Data Overview

      For this project I will utilize data collected and aggregated by the Centers for Medicare Services (CMS) for all CMS certified nursing homes throughout the United States. The data are accessed here: https://data.cms.gov/provider-data/archived-data/nursing-homes. There are approximately 15,000 CMS certified nursing homes throughout the US, so we will have a large geographically diverse population to study.

For this analysis I will be combining two files:

      The provider information file contains general information on currently active nursing homes, including number of certified beds, quality measure scores, staffing and other information used in the Five-Star Rating System. Data are presented as one row per nursing home.

    The second file will contain data aggreated for the Skilled Nursing Facility Quality Reporting Program. CMS aggregates data from the over 15,000 certified nursing homes for a number of quality metrics. They provide these metrics on their Care Compare Website.

      Detailed infromation about the various QRP metrics can be found on CMS technical appendix page: https://www.cms.gov/Medicare/Quality-Initiatives- Patient-Assessment-Instruments/NursingHomeQualityInits/Skilled-Nursing-Facility-Quality-Reporting- Program/SNF-Quality-Reporting-Program-Measures-and-Technical-Information.

      The primary QRP metric of interest for this study will be S_039_01: Percentage of  patients who acquired a healthcare-associated infection during their SNF stay that resulted in hospitalization. This metric is derived entirely from fee-for-service Medicare claims data and relatively new, as it has only been publicly reported since late 2022.  This percentage is shown both as the observed rate and as the risk standardized rate. CMS also reports if this rate is as expected, significantly worse than expected, or significantly better compared to all other nursing homes in that time period.

Data Analysis

      The prposed study aims to assess various nursing home characteristics for association with healthcare-associated infection rates.

      First, we will provide a descriptive analysis of nursing homes across the United States including location, bed size, ownership, average number of residents per day. Specifically of interest, we will also describe variability in overall quality ratings, the healthcare-associated infection quality reporting measure, and measures of staffing (including various levels of nursing care (i.e., RN, LPN, and aide)  both aggregated and separately). 
      Second, we will utilize ANOVA methods to determine which, if any,  average staffing characteristics are associated with the classifications of HAI rates as “better”, “worse”, or “as expected”.  
      Third, we will assess a variety of nursing home characteristics and selected staffing measures for associations with  the risk adjusted rate of HAIs in the nursing home.

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2023-02-25 code/analysis_code/exploratory_analysis.qmd
Project Part 2

---
title: "Exploratory analysis "
author: "Kelly Hatfield"
date: "02/24/2023"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)

#Reminder of variable names
ls(mydata)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Histogram plots for the continuous outcomes.

First, Risk Standardized HAI Rate
```{r}
p1 <- mydata %>% ggplot(aes(x=HAI_039_RS_Rate)) + geom_histogram() 
plot(p1)
figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
ggsave(filename = figure_file, plot=p1)

```

Now Total Nurse Staffing per Resident per Day

```{r}
p2 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_histogram() 
plot(p2)
figure_file = here("results","Adj_Tot_Staff_prpd.png")
ggsave(filename = figure_file, plot=p2) 
```

Now showing relationship between staffing and HAI rates

```{r}
p3 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, y=HAI_039_RS_Rate, color=Staffing.Rating)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results","staff_hai.png")
ggsave(filename = figure_file, plot=p3) 
```

Once more height as function of weight, stratified by ownership type.

```{r}
p4 <- mydata %>% ggplot(aes(x=Ownership.Type, y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","staff_hai_ownership.png")
ggsave(filename = figure_file, plot=p4) 
```

# Notes

Might want to regroup ownership type into fewer categories (e.g., put all government and for-profit together.

### 2023-02-25 code/processing_code/processingfile_v1.qmd
Project Part 2

---
title: "Cleaning Script"
author: "Kelly Hatfield"
date: "2023-02-23"
output: html_document
---

# Processing script

This code contains script to load and process/clean the data.

# Setup

Load needed packages. make sure they are installed.

```{r}

library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library (tidyverse) #for data processing/cleaning
library(readr)
```

# Data loading

```{r}
# path to data
# note the use of the here() package and not absolute paths
provider_location <- here::here("data","raw_data","NH_ProviderInfo_Jan2023.csv")
raw_provider <- read.csv(provider_location)

quality_location <- here::here("data","raw_data","CMS Nursing Home HAI039.csv")
raw_quality <- read_csv(quality_location)
View(raw_quality)
ls(raw_quality)

```

# Check data

Several ways of looking at the data. Lets start wtih a glimpse of the provider data.

```{r}
dplyr::glimpse(raw_provider)

```

# Cleaning

There are a lot of variables and observations in this dataset. We want to subset to just the variables of interest, first.

```{r}
raw_provider1 <- raw_provider[c("Federal.Provider.Number", "Average.Number.of.Residents.per.Day",
                   "Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day",
                   "Staffing.Rating", "Provider.State",   "Ownership.Type")]
skim(raw_provider1)

```

We see a spattering of missing values throughout the dataset with the majority of missing values for total staffing per resident per day. However for a dataset of over 15,000 it is mostly complete. We will leave in the missing values until we select which variables we will incorporate into our model.

Next, we need to explore the quality metric data.

```{r}
#Observe Variable Names
ls(raw_quality)
skim(raw_quality)
```

We know from our data gathering phase that nursing homes that didn't have enough eligible residents for measure calculation would not have a HAI measure calculated. We see 3,209 of those cases are missing here. We will exclude them from our analysis. Also, we can get rid of the unnecessary variables too.

```{r}

raw_quality1 <- raw_quality %>% 
  select (cms_certification_number_ccn, state, HAI_039_Observed, HAI_039_Obs_Rate, HAI_039_Denominator, HAI_039_RS_Rate)

skim(raw_quality1)
```

Now we have no more missing in the quality metrics!

Finally, we need to merge these datasets together. The common identifer variable is the CCN. But we will need to rename the variable from the provider table first. Then, get rid of the nursing homes with no HAI quality measures.

```{r}
raw_provider2 <- raw_provider1 %>% rename(cms_certification_number_ccn="Federal.Provider.Number")

raw_combined <- merge(raw_provider2, raw_quality1, by = "cms_certification_number_ccn")

raw_combined2 <- raw_combined %>% filter(!is.na(HAI_039_Observed))
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- raw_combined2
```

# Save data

```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

### 2023-03-18 code/analysis_code/exploratory_analysis.qmd
Part 3

@@
 #load needed packages. make sure they are installed.
-library(here) #for data loading/saving
-library(dplyr)
+library(here) 
+library (tidyverse)
+
 library(skimr)
+
 library(ggplot2)
@@
 First, Risk Standardized HAI Rate
+
 ```{r}
@@
 ggsave(filename = figure_file, plot=p3)
+
+p3b <- mydata %>% ggplot(aes(x=as.factor(Staffing.Rating), y=HAI_039_RS_Rate)) + geom_boxplot() + geom_jitter(width=0.15, alpha=0.08, color="black", fill="white")
+plot(p3b)
+
+figure_file = here("results","staffrating_hai.png")
+ggsave(filename = figure_file, plot=p3b)
+
 ```
@@
 ```{r}
-p4 <- mydata %>% ggplot(aes(x=Ownership.Type, y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
+p4 <- mydata %>% ggplot(aes(x=as.factor(Ownership.Type), y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
 plot(p4)
@@
 
-# Notes
+# Changing our outcome
 
-Might want to regroup ownership type into fewer categories (e.g., put all government and for-profit together.
+In this plot we categorized our outcome as a 3-level variable: better than the national rate, worse than the national rate or no different than the national rate.
+
+```{r}
+
+p5 <-  mydata %>% ggplot(aes(x=as.factor(COMP_PERF), y=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_boxplot() + geom_jitter(width=0.15, alpha=0.08, color="black", fill="white")
+plot(p5)
+
+figure_file = here("results","staff_hai_compscore.png")
+ggsave(filename = figure_file, plot=p5)
+
+```
+
+We find that while nursing homes who have rates better than the national rate do appear to have higher staffing scores, nursing homes no different and worse than the national rate appear more similar in distribution.
+
+## Modeling Fun
+
+For this section I have decided I am going to assess if staffing rating adds additional information to my model that includes staffing rate.
+For this model we will use the risk adjusted HAI rate as a continuous outcome.
+We will use a linear model.
+We will make three models:
+
+1.  Model with staffing rate and other covariates of interest (Ownership Type, Average Number of Residents per Day, and staffing rate (the continuous value)
+2.  Model with staffing rating and other covariates of interest (Ownership Type, Average Number of Residents per day, and staffing rating (the categorical value)
+3.  Model with both staffing rating and staffing rate + other covariates of interest
+
+```{r}
+#Model 1
+
+model1 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)
+
+summary(model1)
+
+model2 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + as.factor(Staffing.Rating)+ Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)
+
+summary(model2)
+
+model3 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day+ as.factor(Staffing.Rating) + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)
+
+summary(model3)
+
+```
+
+My key takeaways from these three models is that Model 2 and Model 3 have very similar R\^2 fits.
+So they are likely not adding that much information by including both staffing rating and the staffing rate in the same model.
+Also, while less important, we see that staffing rate is not significantly associated with our outcome in this fully adjusted model.
+
+I think currently I am leaning towards choosing model 2 as a our final model for associations.
+This will allow us to draw conclusions about how staffing ratings are associated with healthcare associated infection rates.

### 2023-03-18 code/processing_code/processingfile_v1.qmd
Part 3

@@
                    "Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day",
-                   "Staffing.Rating", "Provider.State",   "Ownership.Type")]
+                   "Staffing.Rating", "Provider.State",   "Ownership.Type", 
+                   "Total.Weighted.Health.Survey.Score", "Number.of.Citations.from.Infection.Control.Inspections")]
 skim(raw_provider1)
@@
 raw_quality1 <- raw_quality %>% 
-  select (cms_certification_number_ccn, state, HAI_039_Observed, HAI_039_Obs_Rate, HAI_039_Denominator, HAI_039_RS_Rate)
+  select (cms_certification_number_ccn, state, HAI_039_Observed, HAI_039_Obs_Rate, HAI_039_Denominator, HAI_039_RS_Rate, COMP_PERF)
 

### 2023-03-18 products/manuscript/Manuscript.qmd
Part 3

@@
 
-Part 1: Project Overview
+The structure below is one possible setup for a data analysis project (including the course project).
+For a manuscript, adjust as needed.
+You don't need to have exactly these sections, but the content covering those sections should be addressed.
+
+This uses MS Word as output format.
+[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
+You can switch to other formats, like html or pdf.
+See [the Quarto documentation](https://quarto.org/) for other formats.
+
+```{r, echo=FALSE, message=FALSE}
+# load a few R packages
+library(here)
+library(knitr)
+```
+
+# Summary/Abstract
 
@@
 
-      The prposed study aims to assess various nursing home characteristics for association with healthcare-associated infection rates.
+The proposed study aims to assess the relationship between overall staffing rates with healthcare-associated infection rates.
 
-      First, we will provide a descriptive analysis of nursing homes across the United States including location, bed size, ownership, average number of residents per day.
-Specifically of interest, we will also describe variability in overall quality ratings, the healthcare-associated infection quality reporting measure, and measures of staffing (including various levels of nursing care (i.e., RN, LPN, and aide)  both aggregated and separately).
-      Second, we will utilize ANOVA methods to determine which, if any,  average staffing characteristics are associated with the classifications of HAI rates as “better”, “worse”, or “as expected”.  
+First, we will provide a descriptive analysis of nursing homes across the United States including two different staffing measures, bed size, ownership, average number of residents per day.
+Second, we will utilize ANOVA methods to determine which, if any, average staffing characteristics are associated with the classifications of HAI rates as "better", "worse", or "as expected".
 Third, we will assess a variety of nursing home characteristics and selected staffing measures for associations with the risk adjusted rate of HAIs in the nursing home.
-
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
-
-# Summary/Abstract
-_Write a summary of your project._
 
@@
 ## General Background Information
-_Provide enough background on your topic that others can understand the why and how of your analysis_
+
+*Provide enough background on your topic that others can understand the why and how of your analysis*
 
 ## Description of data and data source
-_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
+
+*Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section.*
 
 ## Questions/Hypotheses to be addressed
-_State the research questions you plan to answer with this analysis._
+
+*State the research questions you plan to answer with this analysis.*
 
@@
 
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
+*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*
 
 ## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
+
+*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*
 
 ## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
+
+*Write code that reads in the file and cleans it so it's ready for analysis.
 Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
@@
 as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
+All materials, including files that contain code, should be commented well so everyone can follow along.*
 
 ## Statistical analysis
-_Explain anything related to your statistical analyses._
+
+*Explain anything related to your statistical analyses.*
 
@@
 
-_Use a combination of text/tables/figures to explore and describe your data.
+*Use a combination of text/tables/figures to explore and describe your data.
 Show the most important descriptive results here.
 Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
+Even more can be in the R and Quarto files that are part of your project.*
 
@@
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
+*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
 simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
+Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
 
@@
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
+*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
 This might again be code that is best placed in one or several separate R scripts that need to be well documented.
 You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
+Then you load them here.*
 
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
+
+*Summarize what you did, what you found and what it means.*
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
+
+*Discuss what you perceive as strengths and limitations of your analysis.*
 
 ## Conclusions
-_What are the main take-home messages?_
 
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
+*What are the main take-home messages?*
+
+*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
 

### 2023-04-08 code/analysis_code/exploratory_analysis.qmd
More updates

@@
 library(ggplot2)
+
+library(tidymodels)
+library(performance)
 ```
@@
 
-## Modeling Fun
+## Modeling Exercises
+
+First, I have decided to split the dataset into test and training sets.
+
+```{r}
+# Fix the random numbers by setting the seed 
+# This enables the analysis to be reproducible when random numbers are used 
+set.seed(555)
+# Put 3/4 of the data into the training set 
+data_split <- initial_split(mydata, prop = 3/4)
+
+# Create data frames for the two sets:
+train_data <- training(data_split)
+test_data  <- testing(data_split)
+
+```
 
@@
 
+#Applying the model to unseen test data
+
+predict(model_fit2, test_data)
+
+model_aug2\<- augment(model_fit1, test_data)
+
+#pull RS HAI Rate and predictions
+
+temps \<- model_aug2 %\>% pull(HAI_039_RS_Rate)
+
+pred \<- model_aug2 %\>% pull(.pred)
+
+#run rmse
+
+rmse_vec(temps, pred)
+
 ```{r}
@@
 
-model1 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)
+recipe_1 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)
 
-summary(model1)
+recipe_2 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + as.factor(Staffing.Rating), data= train_data)
 
-model2 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + as.factor(Staffing.Rating)+ Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)
+recipe_3 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + as.factor(Staffing.Rating), data= train_data)
 
-summary(model2)
+#model: linear regression using GLM engine
+ln_mod <- linear_reg() %>% set_engine ("lm")
 
-model3 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day+ as.factor(Staffing.Rating) + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)
+#workflow: tying together linear model and recipe of chocie
+wflow1 <- workflow() %>% add_model(ln_mod) %>% add_recipe(recipe_1)
+wflow2 <- workflow() %>% add_model(ln_mod) %>% add_recipe(recipe_2)
+wflow3 <- workflow() %>% add_model(ln_mod) %>% add_recipe(recipe_3)
 
-summary(model3)
+#Fitting the model to train dataset
+model_fit1 <- wflow1 %>%  fit(data = train_data)
+model_fit2 <- wflow2 %>%  fit(data = train_data)
+model_fit3 <- wflow3 %>%  fit(data = train_data)
+
+#Looking at model output;
+model_fit1 %>% extract_fit_parsnip() %>%  tidy()
+model_fit2 %>% extract_fit_parsnip() %>%  tidy()
+model_fit3 %>% extract_fit_parsnip() %>%  tidy()
+
+#Applying the model to unseen test data
+predict(model_fit2, test_data)
+model_aug2<- augment(model_fit1, test_data)
+
+#pull RS HAI Rate and predictions
+temps <- model_aug2 %>% pull(HAI_039_RS_Rate)
+pred <- model_aug2 %>% pull(.pred)
+
+#run rmse
+rmse_vec(temps, pred)
+
+#Selecting model 2 as final model
+
+#Running on test data
 

### 2023-04-08 products/manuscript/Manuscript.qmd
More updates

@@
 
-The structure below is one possible setup for a data analysis project (including the course project).
-For a manuscript, adjust as needed.
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
 ```{r, echo=FALSE, message=FALSE}
@@
 
-# Summary/Abstract
+# Summary
 
@@
 
-*Provide enough background on your topic that others can understand the why and how of your analysis*
+Over 15,000 skilled nursing facilities (SNFs) and nursing homes provide residential long term care including both personal and medical services to over a million patients throughout the United States.
+Due to advanced age and medical complexity of residents in the nursing home setting, this population is at increased risk for significant morbidity and mortality.
+Previous estimates show that approximately 12% of NH residents had an infection at a given time; most commonly urinary tract infection (3.0--5.2%), pneumonia (2.2--4.4%), and cellulitis (1.6--2.0%).
+In the COVID-19 pandemic, nearly all nursing homes experienced outbreaks of COVID-19, and nursing home residents had disproportionately high mrobdiity and mortality compared to community dwelling adults.
+Nursing homes were mentioned as a specific target for reducing the morbidity and mortality related to healthcare associated infections in the 2009 *National Action Plan to Prevent Health Care-Associated Infections: Roadmap to Elimination* (HAI National Action Plan) and subsequent updates published in 2013 and 2018. Most healthcare associated infections (HAIs) are thought to be preventable as they result from the processes and structures related to poor care, including poor adherence to protocols and guidelines.
 
-## Description of data and data source
+The Centers for Medicare and Medicaid Services (CMS) reports measures of nursing home quality on their Care Compare website (https://www.medicare.gov/care-compare/).
+These measures of care include ownership details, health inspection ratings, staffing ratings, and quality measures.
+Data are searchable and comparable on the website for all participating nursing homes.
+Aggregate datasets are also available for download.
 
-*Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section.*
+Federal law requires all nursing homes to provide enough staff to safely care for residents. Nursing home reports daily staffing hours to Medicare in the payroll based journal.
+Using this information, Medicare calculates a ratio of staffing hours per resident day, the percent of nurse staff that stop working at the facility (turnover), and the number of administrators who have left the facility within a given year and reports the data publicly.
+CMS also creates a staffing star measure that is adjusted based off the needs of the residents in each home and is compiled using aggregated metrics about the RN staffing levels, total staffing hours, weekend staffing hours, total nurse turnover, RN turnover, and administrator turnover.
+(https://www.medicare.gov/care-compare/resources/nursing-home/staffing)
 
-## Questions/Hypotheses to be addressed
+CMS also calculates quality measures star rating.
+This assess different measures of nursing home quality.
+For this report we are most interested in a new quality measure *Skilled Nursing Facility Healthcare-Associated Infections (HAI) Requiring Hospitalization* that became reported for the first time in October 2022, with data collected throughout FY 2019.
+The HAI requiring hospitalization metric estimates the risk-standardized rate of HAIs that are acquired during SNF care andresult in hospitalization.
+The measure is calculated on Medicare claims data from fee-for-service beneficiaries residing within a nursing home.
+The measure requires no additional reporting from nursing home staff, and relies entirely on administrative claim data and is based on principal diagnosis on the Medicare inpatient (IP) claims of SNF residents.
 
-*State the research questions you plan to answer with this analysis.*
+## Study Objectives
+
+We aim to assess the association of nursing home staff star ratings with the healthcare associated infection rate in participating nursing homes.
+
+##
 
@@
 
-*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*
+*Data were accessed from the*
 
@@
 
-*Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project.*
+From all data, we identified 10,502 nursing homes with complete data to include in our analysis. Complete description of the cohort included in the study in the Supplementary Files.
 

### 2023-04-08 products/manuscript/Manuscript.qmd
Updated mansucript

@@
 
-Data Overview
-
-For this project I will utilize data collected and aggregated by the Centers for Medicare Services (CMS) for all CMS certified nursing homes throughout the United States.
-The data are accessed here: https://data.cms.gov/provider-data/archived-data/nursing-homes.
-There are approximately 15,000 CMS certified nursing homes throughout the US, so we will have a large geographically diverse population to study.
-
-For this analysis I will be combining two files:
-
-The provider information file contains general information on currently active nursing homes, including number of certified beds, quality measure scores, staffing and other information used in the Five-Star Rating System.
-Data are presented as one row per nursing home.
-
-The second file will contain data aggreated for the Skilled Nursing Facility Quality Reporting Program.
-CMS aggregates data from the over 15,000 certified nursing homes for a number of quality metrics.
-They provide these metrics on their Care Compare Website.
-
-Detailed infromation about the various QRP metrics can be found on CMS technical appendix page: https://www.cms.gov/Medicare/Quality-Initiatives- Patient-Assessment-Instruments/NursingHomeQualityInits/Skilled-Nursing-Facility-Quality-Reporting- Program/SNF-Quality-Reporting-Program-Measures-and-Technical-Information.
-
-The primary QRP metric of interest for this study will be S_039_01: Percentage of patients who acquired a healthcare-associated infection during their SNF stay that resulted in hospitalization.
-This metric is derived entirely from fee-for-service Medicare claims data and relatively new, as it has only been publicly reported since late 2022.
-This percentage is shown both as the observed rate and as the risk standardized rate.
-CMS also reports if this rate is as expected, significantly worse than expected, or significantly better compared to all other nursing homes in that time period.
-
-Data Analysis
-
-The proposed study aims to assess the relationship between overall staffing rates with healthcare-associated infection rates.
-
-First, we will provide a descriptive analysis of nursing homes across the United States including two different staffing measures, bed size, ownership, average number of residents per day.
-Second, we will utilize ANOVA methods to determine which, if any, average staffing characteristics are associated with the classifications of HAI rates as "better", "worse", or "as expected".
-Third, we will assess a variety of nursing home characteristics and selected staffing measures for associations with the risk adjusted rate of HAIs in the nursing home.
+The proposed study aims to assess the relationship between overall staffing rates with healthcare-associated infection rates in nursing homes throughout the United States.
 
@@
 Due to advanced age and medical complexity of residents in the nursing home setting, this population is at increased risk for significant morbidity and mortality.
-Previous estimates show that approximately 12% of NH residents had an infection at a given time; most commonly urinary tract infection (3.0--5.2%), pneumonia (2.2--4.4%), and cellulitis (1.6--2.0%).
-In the COVID-19 pandemic, nearly all nursing homes experienced outbreaks of COVID-19, and nursing home residents had disproportionately high mrobdiity and mortality compared to community dwelling adults.
+Previous estimates show that approximately 12% of NH residents had an infection at a given time [@thompson2020epidemiology]; most commonly urinary tract infection (3.0--5.2%), pneumonia (2.2--4.4%), and cellulitis (1.6--2.0%).
+In the COVID-19 pandemic, nearly all nursing homes experienced outbreaks of COVID-19, and nursing home residents had disproportionately high morbidity and mortality compared to community dwelling adults.
 Nursing homes were mentioned as a specific target for reducing the morbidity and mortality related to healthcare associated infections in the 2009 *National Action Plan to Prevent Health Care-Associated Infections: Roadmap to Elimination* (HAI National Action Plan) and subsequent updates published in 2013 and 2018. Most healthcare associated infections (HAIs) are thought to be preventable as they result from the processes and structures related to poor care, including poor adherence to protocols and guidelines.
@@
 Data are searchable and comparable on the website for all participating nursing homes.
-Aggregate datasets are also available for download.
+Aggregate data sets are also available for download.
 
@@
 CMS also creates a staffing star measure that is adjusted based off the needs of the residents in each home and is compiled using aggregated metrics about the RN staffing levels, total staffing hours, weekend staffing hours, total nurse turnover, RN turnover, and administrator turnover.
-(https://www.medicare.gov/care-compare/resources/nursing-home/staffing)
+(https://www.medicare.gov/care-compare/resources/nursing-home/staffing).
 
@@
 For this report we are most interested in a new quality measure *Skilled Nursing Facility Healthcare-Associated Infections (HAI) Requiring Hospitalization* that became reported for the first time in October 2022, with data collected throughout FY 2019.
-The HAI requiring hospitalization metric estimates the risk-standardized rate of HAIs that are acquired during SNF care andresult in hospitalization.
+The HAI requiring hospitalization metric estimates the risk-standardized rate of HAIs that are acquired during SNF care and result in hospitalization.
 The measure is calculated on Medicare claims data from fee-for-service beneficiaries residing within a nursing home.
@@
 
+Previous studies have shown that certain staffing metrics are associated with lower quality metrics in US nursing homes [@mukamel2023association, @mukamel2022daily].
+However, they were limited because they did not assess healthcare associated infections requiring hospitalization.
 We aim to assess the association of nursing home staff star ratings with the healthcare associated infection rate in participating nursing homes.
-
-##
-
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]
 
@@
 
-*Data were accessed from the*
+## Data acquisition
 
-## Data aquisition
+For this project I will utilize data collected and aggregated by the Centers for Medicare Services (CMS) for all CMS certified nursing homes throughout the United States.
+The data are accessed here: <https://data.cms.gov/provider-data/archived-data/nursing-homes>.
+There are approximately 15,000 CMS certified nursing homes throughout the US, so we will have a large geographically diverse population to study.
 
-*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*
+First, the provider information file contains general information on currently active nursing homes, including number of certified beds, quality measure scores, staffing and other information used in the Five-Star Rating System.
+Data are presented as one row per nursing home.
+The nurse staffing metrics in this data set include reported and adjusted measures of nurse aide staffing hours per resident per day, licensed professional nurse (LPN) staffing hours per resident per day, registered nurse (RN) staffing hours per resident per day, total nurse staffing hours per resident per day, and total number of nursing staffing hours per resident per day on the weekend.
+The data also includes turnover data including RN turnover, total staff turn over, and administrator turnover.
 
-## Data import and cleaning
+Second, we addend the provider information with aggregated data reported as part ofthe Skilled Nursing Facility Quality Reporting Program.
+CMS aggregates data from the over 15,000 certified nursing homes for a number of quality metrics.
+They provide these metrics on their Care Compare site referenced above.
 
-*Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along.*
+The primary QRP metric of interest for this study is titled S_039_01: Percentage of patients who acquired a healthcare-associated infection during their SNF stay that resulted in hospitalization.
+This percentage is shown both as the observed rate and as a risk standardized rate.
+The HAI requiring hospitalization metric is also given as a categorical variable, where the nursing home is compared to all other nursing homes in that time period and categorized as average, significantly worse than average, or significantly better than average.
+
+## Data cleaning
+
+Data from both data sets will be merged together using a unique facility identifier, the CMS Certification Number (CCN).
+We will subset our data set to include only nursing homes with complete provider information and HAI requiring hospitalization metrics.
+Nursing homes with missing data will be excluded from analysis.
 
@@
 
-*Explain anything related to your statistical analyses.*
+First, we describe the distribution of adjusted average total staffing and risk standardized rate of HAIs requiring hospitalization by nursing home.
+We also describe the frequency and percent of the categorical staffing rating and HAI categorization (better, worse, or same as average).
+Second, we consider bivariate associations of the risk standardized infection rate by the staffing rating variable.
+
+Third, we create multivariable models that assess the relationship between the risk standardized HAI rate (outcome) and the total average staffing per resident per day and staffing rating, while adjusting for important nursing home characteristics as identified through the literature including: ownership, number of health violation, health inspection score, and average number of residents per day.
+For our multivariable models, we aim to determine if the total average staffing per resident per day and staffing rating are independently associated with the outcome.
+
+To create our multivariable models and assess their ability for prediction, we first divided our full data cohort into a training (70% of nursing homes) and test data set (30% of nursing homes).
+We used our training data to determine covariates included in our final model.
+Then, we applied our final model to our test data subset.
+We compared parameter estimates and conclusions, to ensure that the model fits similarly in both subsets of data.
 
@@
 From all data, we identified 10,502 nursing homes with complete data to include in our analysis. Complete description of the cohort included in the study in the Supplementary Files.
+
+Figure 1
 

### 2023-04-21 code/analysis_code/analysis for paper.qmd
Peer Review Push 1

---
title: "Paper analysis "
author: "Kelly Hatfield"
date: "04/23/2023"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) 
library (tidyverse)

library(skimr)

library(ggplot2)

library(tidymodels)
library(performance)
library(stringr)
library(dplyr)

```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)

#Reminder of variable names
ls(mydata)
```

## Results

### Descriptive stats: staff and HAI rate

f2

figure2 = here("results","Staff_prpd_distribution.png")

ggsave(filename = figure2, plot=f2)

```{r}

mean(mydata$HAI_039_RS_Rate)

mytitlef1 <- "Figure: Distribution of adjusted rate of HAIs requring hospitalization by nursing home"

p1 <- mydata %>% ggplot(aes(x=HAI_039_RS_Rate)) + geom_histogram() + labs(
    x = "Adjusted rate of HAIs requring hospitalization",              # x axis title
    y = "Count of Nursing Homes",   # y axis title
    title=  str_wrap(mytitlef1, 50) ,      # main title of figure
    colour = "Continent"      # title of legend
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(face = "bold"))

figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
ggsave(filename = figure1, plot=p1)

mydata %>%
  group_by(COMP_PERF) %>% summarise( n = n(), percent=n()/sum(n()),
                                     mean = mean(HAI_039_RS_Rate))

vp <- ggplot(mydata, aes(x=COMP_PERF, y=HAI_039_RS_Rate)) + labs(x = " ", y = "Adjusted rate of HAIs requring hospitalization", title=  str_wrap(mytitlef1, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
vp

mytitlef2 <- "Figure: Distribution of total staff per resident per day by nursing home"

f2 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_histogram() + labs(y = "Count of Nursing Homes ", x = "Total Staffing per Resident per Day", title=  str_wrap(mytitlef2, 50))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(face = "bold"))

f2
figure2 = here("results","Staff_prpd_distribution.png")
ggsave(filename = figure2, plot=f2)

vp2 <- ggplot(mydata, aes(x=as.factor(Staffing.Rating), y=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + labs(x = " ", y = "Total Nurse Staffing Per Resident Per Day", title=  str_wrap(mytitlef1, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
vp2
figure2b = here("results","Staff_prpd_distribution_group.png")
ggsave(filename = figure2b, plot=vp2)

mydata %>%
  group_by(Staffing.Rating) %>% summarise( n = n(), percent=n()/sum(n()),
                                     mean = mean(Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day))

```

## Bivariate Analysis

```{r}
library(ggpubr)

mytitlef3 <- "Figure: Relationship of total staff per resident per day by nursing home with Adjusted HAI Rate by nursing home"
sp <- ggplot(mydata, aes(x= Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, y = HAI_039_RS_Rate)) + geom_point() + labs(y = "Adjusted HAI Rate ", x = "Total Nurse Staffing Per Resident Per Day", title=  str_wrap(mytitlef1, 75)) + geom_smooth(method=lm, se=FALSE)  + theme(plot.title = element_text(hjust = 0.5), panel.background = element_blank(), axis.line = element_line(color="black"), axis.line.x = element_line(color="black")) + stat_cor(method="pearson", label.x=0, label.y=20)

sp

figure3 = here("results","correlation.png")
ggsave(filename = figure3, plot=sp)

mydata %>%
  group_by(COMP_PERF) %>% summarise( n = n(), percent=n()/sum(n()),
                                     mean = mean(Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, na.rm=T))

mytitlef4a <- "Figure: Relationship of total staff per resident per day by nursing home with HAI Rating by nursing home"

vp4a <- ggplot(mydata, aes(x=as.factor(COMP_PERF), y=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + labs(x = " ", y = "Total Nurse Staffing Per Resident Per Day", title=  str_wrap(mytitlef4a, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
figure4 = here("results","staffingrate_HAIgroup.png")
ggsave(filename = figure4, plot=vp4a)

###### Save for later ######
mytitlef4b <- "Figure: Relationship of adjusted HAI rate with Staffing Rating by nursing home"
vp4b <- ggplot(mydata, aes(x=as.factor(Staffing.Rating), y=HAI_039_RS_Rate)) + labs(x = " ", y = "Adjusted HAI Rate", title=  str_wrap(mytitlef4b, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
vp4b

```

## Models

```{r}
# This enables the analysis to be reproducible when random numbers are used 
set.seed(555)

mydata$Staffing.Rating.F <- as.factor(mydata$Staffing.Rating)

# Put 3/4 of the data into the training set 
data_split <- initial_split(mydata, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)

recipe_2 <- recipe(HAI_039_RS_Rate ~ Staffing.Rating  + Ownership.Type + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)

#model: linear regression using GLM engine
ln_mod <- linear_reg() %>% set_engine ("lm")

#workflow: tying together linear model and recipe of chocie
wflow2 <- workflow() %>% add_model(ln_mod) %>% add_recipe(recipe_2)

#Fitting the model to train dataset
model_fit2 <- wflow2 %>%  fit(data = train_data)

#Looking at model output;
model_fit2 %>% extract_fit_parsnip() %>%  tidy()

#Applying the model to unseen test data
predict(model_fit2, test_data)
model_aug2<- augment(model_fit2, test_data)

#pull RS HAI Rate and predictions
temps <- model_aug2 %>% pull(HAI_039_RS_Rate)
pred <- model_aug2 %>% pull(.pred)

#run rmse
rmse_vec(temps, pred)

#Looking at model output on all of the data
wflow2 %>%  fit(data = mydata) %>% extract_fit_parsnip() %>%  tidy()
```

# Appendix

## Table 1: Summary of Findings

Showing a bit of code to produce and save a summary table.

```{r}

#Appendix Tabe 1: Make a table of all the distributions of all variables included in the analysis univariate

summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)

```

##

### 2023-04-21 code/analysis_code/exploratory_analysis.qmd
Peer Review Push 1

@@
 library(performance)
+library(stringr)
 ```
@@
 ```{r}
-p2 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_histogram() 
+
+mytitle <- "Figure: Distribution of adjusted total nursing staffing hours per resident per day"
+
+p2 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_histogram() + labs(
+    x = "Adjusted total nursing staffing hours per resident per day",              # x axis title
+    y = "Count of Nursing Homes",   # y axis title
+    title=  str_wrap(mytitle, 60) ,      # main title of figure
+    colour = "Continent"      # title of legend
+  ) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1),
+        plot.title = element_text(face = "bold"))
+
 plot(p2)
@@
 predict(model_fit2, test_data)
-model_aug2<- augment(model_fit1, test_data)
+model_aug2<- augment(model_fit2, test_data)
 
@@
 Also, while less important, we see that staffing rate is not significantly associated with our outcome in this fully adjusted model.
-
-I think currently I am leaning towards choosing model 2 as a our final model for associations.
-This will allow us to draw conclusions about how staffing ratings are associated with healthcare associated infection rates.

### 2023-04-21 products/manuscript/Manuscript.qmd
Peer Review Push 1

@@
 
-## General Background Information
+## Background
 
@@
 
-The Centers for Medicare and Medicaid Services (CMS) reports measures of nursing home quality on their Care Compare website (https://www.medicare.gov/care-compare/).
+The Centers for Medicare and Medicaid Services (CMS) reports measures of nursing home quality on their [Care Compare website](https://www.medicare.gov/care-compare/ "https://www.medicare.gov/care-compare/").
 These measures of care include ownership details, health inspection ratings, staffing ratings, and quality measures.
@@
 Using this information, Medicare calculates a ratio of staffing hours per resident day, the percent of nurse staff that stop working at the facility (turnover), and the number of administrators who have left the facility within a given year and reports the data publicly.
-CMS also creates a staffing star measure that is adjusted based off the needs of the residents in each home and is compiled using aggregated metrics about the RN staffing levels, total staffing hours, weekend staffing hours, total nurse turnover, RN turnover, and administrator turnover.
-(https://www.medicare.gov/care-compare/resources/nursing-home/staffing).
+CMS also creates a [staffing star measure](https://www.medicare.gov/care-compare/resources/nursing-home/staffing) that is adjusted based off the needs of the residents in each home and is compiled using aggregated metrics about the RN staffing levels, total staffing hours, weekend staffing hours, total nurse turnover, RN turnover, and administrator turnover.
 
@@
 
-Second, we addend the provider information with aggregated data reported as part ofthe Skilled Nursing Facility Quality Reporting Program.
+Second, we addend the provider information with aggregated data reported as part of the Skilled Nursing Facility Quality Reporting Program.
 CMS aggregates data from the over 15,000 certified nursing homes for a number of quality metrics.
@@
 We also describe the frequency and percent of the categorical staffing rating and HAI categorization (better, worse, or same as average).
+
 Second, we consider bivariate associations of the risk standardized infection rate by the staffing rating variable.
@@
 
-Figure 1
+Distributions of all variables are included in the Appendix.
 
-@tbl-summarytable shows a summary of the data.
+## Statistical analysis
 
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
+### Univariate Analysis
+
+Overall, 408 (4%) of nursing homes had an adjusted rate worse than average (mean rate: 12.3), and 149 (1%) had an adjusted rate categorized as better than average (mean rate: 4.4).
+The mean adjusted rate of HAIs requiring hospitalization for 9,945 nursing homes categorized no different than average was 7.5.
+The adjusted HAI rates are visualized in Figure 1.
+
+```{r}
+#| label: fig-result1
+#| fig-cap: "Risk Standardized Rate of HAI by Nursing Home"
+#| echo: FALSE
+knitr::include_graphics(here("results","RiskStandardized_HAI_Rate_distribution.png"))
+
+```
+
+Overall, 3,117 (30%) of nursing homes had a staffing rating of 1, and no data available for the total staffing per resident per day.
+Another 154 (1%) of nursing homes were missing staffing rating and staffing measures.
+The distribution stratified by staffing rating is shown in Figure 2.
+The average adjusted total staffing rate was 3.36 for for staffing rating of 2 stars, 3.76 for staffing rating of 3 stars, 4.01 for staffing rating of 4 stars, and 5.30 for a staffing rating of 5 stars.
+
+```{r}
+
+#| label: fig-result2
+#| fig-cap: "Total Staff Per Resident Per Day by Nursing Home"
+#| echo: FALSE
+knitr::include_graphics(here("results","Staff_prpd_distribution.png"))
+
+```
+
+```{r}
+
+#| label: fig-result2b
+#| fig-cap: "Total Staff Per Resident Per Day Stratified by Nursing Home"
+#| echo: FALSE
+knitr::include_graphics(here("results","Staff_prpd_distribution_group.png"))
+
+```
+
+### Bivariate Analysis
+
+The correlation between adjusted average total staffing and the adjusted HAI rate was -0.2 (Figure 3), indicating that the HAI rate decreased as the total staffing measure increased.
+
+```{r}
+#| label: fig-result3
+#| fig-cap: "Staffing rate and Adjusted HAI Rate"
+#| echo: FALSE
+knitr::include_graphics(here("results","correlation.png"))
+
+```
+
+Nursing homes categorized as better than average had a mean adjusted average total staffing rate of 4.79, compared to 3.65 for nursing homes categorized as average, and 3.51 for nursing homes categorized as worse than average (Figure 4).
+
+```{r}
+#| label: fig-result4
+#| fig-cap: "Staffing rate stratified by HAI categorization"
+#| echo: FALSE
+knitr::include_graphics(here("results","staffingrate_HAIgroup.png"))
+```
+
+### Multivariable Models
+
+For our training data, we selected 75% of the data which represented XX nursing homes.
+We compared three linear regression models with the outcome of risk adjusted HAI rate.
+The predictor of interest in the first model was the the total average staffing per resident per day and staffing **rating,** a numeric score from 1 - 5 stars.
+The predictor of interest for the second model was the average total staffing per resident per day score.
+For the third model we considered both the staffing rating and the average total staffing measure as the predictors of interest.
+
+We compared the three models using RMSE to determine which model best fit our data in the training dataset. The comparable RMSE rates are shown in Table 2.
+
+***#Add R Code here to show Table 2 (RMSE for different model choices 1,2,3)***
+
+We selected model 1, containing only the staffing rating as our best fitting model.When we applied model 1 to our test data we found that the predicted and observed values were fairly similar (Figure 5) and the RMSE was 1.79.
+
+***#Add R Code here to show Figure 5***
+
+For model 1, we found that in our training dataset the parameter estimate for staffing rating was -0.2.
+This can be interpreted as for each increase in staffing rating, the adjusted HAI rate decreases by 0.2, while adjusting for other nursing home characteristics.
+When we ran model 2 on our full dataset, we also got a parameter estimate of -0.2.
+This gives us confidence that our conclusions from model 1 are robust.
 
@@
 
-## Basic statistical analysis
-
-*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-knitr::include_graphics(here("results","height_weight_stratified.png"))
-```
-
-## Full analysis
-
-*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here.*
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
-
 {{< pagebreak >}}
@@
 
-*Summarize what you did, what you found and what it means.*
+This analysis demonstrates an important relationship between total staffing measures and risk adjusted healthcare associated infection (HAI) rates.
+Using over 10,000 nursing homes located throughout the United States, we have found that adjusted HAI rates are lower in nursing homes that have higher measures of total staffing or better staffing rating scores.
+
+This provides additional evidence to support the need of increased nursing staff in a nursing home setting to improve patient safety and quality of care for nursing home residents.
+
+Because increased HAIs have substantial morbidity and mortality and additional healthcare costs, policy makers, payers, and consumers of healthcare should create incentive programs for nursing home administrators to increase their total staffing rates.
+By investing in increased staffing in the nursing home setting you may reduce HAIs and prevent unnecessary harm, prevent hospitalizations, and lower total healthcare costs.
 
@@
 
-*Discuss what you perceive as strengths and limitations of your analysis.*
+This study has strength that it is an assessment of a large number (over 10,000) nursing homes. It includes data from across the United States and represents the diverse landscape of long term care.
+
+Assessment of quality in nursing home settings is limited by the quality of the measures used to assess.
+Our analysis is limited in that it excludes nursing homes that were not able to ascertain a risk adjusted HAI rate.
+Missing data is mostly due to a small number of fee-for-service beneficiaries in that nursing home for the year, and may not be missing at random.
 
@@
 
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
-
 {{< pagebreak >}}

### 2023-04-28 reviews/kailin_chen.qmd
project peer evaluation

---
title: Project Review Template 
author: Kailin (Kai) Chen
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---
# Overview

**Title of Project**: "Evaluating Staffing Characteristics, Healthcare-associated Infections, and Provider Characteristics in U.S. Nursing Homes, 2021"

**Name of Project Author(s)**: Kelly Hatfield

**Name of Project Reviewer**: Kailin (Kai) Chen

# Specific Project Content Evaluation

## Background, Context, and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The project context is well-described with a comprehensive background referencing the target population, served by residential long-term care through skilled nursing facilities and nursing homes, being at risk for significant morbidity and mortality by infections. It is well-placed in the context of existing work, as this project points out previous research does not link together staffing metrics with healthcare-associated infections, which the author now intends to explain through this data analysis project.

### Summary Assessment
*   Strong contextualization and motivation

## Question Description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments
No specific question or hypothesis was present in this project, only that the author aims "to assess the association of nursing home staff star ratings with the healthcare associated infection rate in participating nursing homes". It should be made clear in the future what the question (presumably "What is the association of nursing home staff star ratings with the healthcare-associated infection rate in participating nursing homes?") and hypothesis are.

### Summary Assessment
*   Question/hypothesis unclear

## Data Description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
Overall, the data is well-described and has a source provided (Centers for Medicare Services). No meta-information, other than a data dictionary, was given to further explain the data. It would be preferable for the author to explicitly mention that *both* datasets used are found from the linked website's downloadable zip files, as well as which zip file was downloaded specifically.

### Summary Assessment
*   Source and overall structure of data well explained

## Data Wrangling and Exploratory Analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments
Data cleaning and processing steps seem very straightforward and reasonable, though from only viewing the the manuscript, no detail is given as to *why* these steps were chosen. Meaningful exploratory analysis results are shown in the manuscript regarding general relationships between nursing staff and healthcare-associated infections.

### Summary Assessment
*   Some weaknesses in wrangling and exploratory analysis

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments
The analysis methods used (bivariate analysis and multivariate models) were appropriate for the data, and analysis was done properly. Different steps of the analysis were completed using the most suitable methodologies (ex: variable selection based on existing literature and model evaluation based on RMSE).

### Summary Assessment
*   Strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments
Overall, results feel unpolished. Most figures were easy to read, but one graph (Figure 1c?) appeared incomplete without an x-axis title. Labeling of figures as 1, 2, 3, etc. were not consistent. Additionally, the author had written "we selected 75% of the data which represented XX nursing homes"; XX seems to be a placeholder for a number. The table added into the manuscript could be presented better as well. I would advise the author to review this section for completeness and appearance quality.

### Summary Assessment
*   Results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
Study findings and strengths/limitations are interpreted and discussed properly, but the conclusion section is blank.

### Summary Assessment
*   Minor parts wrong, missing, or unclear

## Further Comments
As previously mentioned, I would suggest reviewing your exploratory and statistical analyses for completeness and appearance in addition to looking over your entire manuscript for completeness.

# Overall Project Content Evaluation

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments
The project is well-structured overall, with pertinent files in appropriately-labeled folders that indicate workflow. However, not all template files were removed, such as the `slides` subdirectory in `products` and `supplement` under the `manuscript` subdirectory. It is also a good idea to look over `analysis for paper.qmd` and `exploratory_analysis.qmd` and tidy up those sections.

### Summary Assessment
*   Mostly clear but some confusing parts (e.g. useless files, things in the wrong folders)

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Across the different qmd files for this analysis, the project is overall well-documented, and I was able to understand the majority of steps for the analysis.

### Summary Assessment
*   Fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

Portions of `exploratory_analysis.qmd` and `analysis for paper.qmd` are not reproducible without manual intervention.

### Summary Assessment
*   Small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The study overall had a decent level of thoroughness, but the statistical analysis focused on standard regression modeling (even for the machine learning component). I believe this author's results addressed what they were looking for, even if a specific question/hypothesis was never made clear,

### Summary Assessment
*   Decent level of thoroughness

## Further Comments
A good way to check reproducibility is to completely clear the environment and run all qmd files in the way you intend them to be run. Double-check that any template remnants are gone in your review.

### 2023-04-28 MADA_Reviews/MADA_Project_Review_(Perez,K).qmd
Project Peer Review (Perez, K)

---
title: Project Review Template 
author: Kimberly Perez
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Evaluating Staffing Characteristics, Healthcare-associated Infections, and Provider Characteristics in U.S. Nursing Homes, 2021

Name of project author(s): Kelly Hatfield

Name of project reviewer: Kimberly Perez

## Background, Context and Motivation

Very thorough background. I was able to follow along easily and while I do not know much about this topic, you draw in the reader with this section. Well-written with minor, if any, grammatical errors.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

Your question was succinct and even provided a bit more background information on topic. Further clarifies the importance of your study/analysis. Consider including a hypothesis in conjunction with the objectives

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
Very clear explanation as to how the data was collected and additions made to the dataset. Your inclusion of the title of files is helpful when following along.

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Great job with this section. I was able to execute the wrangling process with no issues! Nice work merging unique identifiers. I would consider adding some more notes to some sections as you walk through the process, however, if time is an issue, your current structure works! For your exploratory analysis, I received an error on line 175 (error in `line_check()`, I was not able to run the remaining code.

### Summary assessment
* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

### Feedback and Comments

I received an error when attempting to run your analysis on line 67- Error in force(filename) : object ‘figure1’ not found. Manual intervention allowed me to run the remained of the code. The analysis seems to be sufficient based on the data. Consider including some more information on your outputs

### Summary assessment
* strong and reasonable analysis

## Presentation

### Feedback and Comments

Overall, nice work with the figures and figure captions. Some coding exists in the manuscript but otherwise all figures look good. If you have time, you may want to consider editing table 1 as it is currently difficult to read and spans a few pages.

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

I thought this area was really well written, succinct, and really drove home the point of your findings. I think expanding a bit upon the strengths and weaknesses and wrapping up with a short conclusion will really wrap-up your manuscript nicely!

### Summary assessment
* strong, complete and clear discussion

## Further comments

Nice job, Kelly! I think you are on your way to a great product!

# Overall project content evaluation

## Structure

Overall, your project is nicely structured. The majority of your readme files are updated with information making your r script easy to follow. Some “junk” folders remain, specifically the “slides” and “supplement” folders found within the products folder.

### Summary assessment
* well structured

## Documentation

I think there is some room, specifically with your analysis, to include a bit more documentation (see above) but overall, you have items properly documented.

### Summary assessment
* decently documented with some gaps

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

Overall, solid work. Some areas (mentioned above) produced errors but nothing major, with manual intervention, I was able to run the remaining code!

### Summary assessment
* small parts not reproducible or required manual intervention

## Thoroughness

### Feedback and Comments
Yes! I liked your inclusion of other models in your analysis script and would encourage you to add whatever is not used in your supplemental material qmd. Overall, I feel like your manuscript is thorough and will likely be very thorough for the final submission.

### Summary assessment
* decent level of thoroughness

## Further comments

Great work, Kelly! I think you will have a great product with a bit of polishing! Please let me know if you have any questions or concerns moving forward.

### 2023-05-05 code/analysis_code/analysis_paper.qmd
Project Analysis Thursday PM

--- 2023-04-21 code/analysis_code/analysis for paper.qmd
@@
 
-f2
+```{r}
 
-figure2 = here("results","Staff_prpd_distribution.png")
-
-ggsave(filename = figure2, plot=f2)
-
-```{r}
+mydata$Staffing.Rating.F <- as.factor(mydata$Staffing.Rating)
+mydata$owner <- substr(mydata$Ownership.Type, 1, 10)
 
@@
 figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
-ggsave(filename = figure1, plot=p1)
+ggsave(filename = figure_file, plot=p1)
 
@@
 
-###### Save for later ######
-mytitlef4b <- "Figure: Relationship of adjusted HAI rate with Staffing Rating by nursing home"
-vp4b <- ggplot(mydata, aes(x=as.factor(Staffing.Rating), y=HAI_039_RS_Rate)) + labs(x = " ", y = "Adjusted HAI Rate", title=  str_wrap(mytitlef4b, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
-               geom = "point",
-               color = "red")
-vp4b
+vp4a
 
@@
 
-mydata$Staffing.Rating.F <- as.factor(mydata$Staffing.Rating)
-
 # Put 3/4 of the data into the training set 
@@
 
-recipe_2 <- recipe(HAI_039_RS_Rate ~ Staffing.Rating  + Ownership.Type + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)
+recipe_2 <- recipe(HAI_039_RS_Rate ~ Staffing.Rating  + owner + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)
 
@@
 
+mytitlescat <- "Observed and Predicted HAI Rates from Test Model"
+scatter_model <- ggplot(model_aug2, aes(x=.pred, y=HAI_039_RS_Rate)) + labs(x = "Predicted Rate ", y = "Adjusted HAI Rate", title=  str_wrap(mytitlescat, 75))+ geom_point() +geom_abline() + stat_cor(method="pearson", label.x=6, label.y=20)
+scatter_model
+
+figure5 = here("results","scatter_model.png")
+ggsave(filename = figure5, plot=scatter_model)
+
 #Looking at model output on all of the data
-wflow2 %>%  fit(data = mydata) %>% extract_fit_parsnip() %>%  tidy()
+table2save<- wflow2 %>%  fit(data = mydata) %>% extract_fit_parsnip() %>%  tidy()
+table2 = here("results", "table2.RDS")
+saveRDS(table2save, file=table2) 
 ```

### 2023-05-05 code/analysis_code/exploratory_analysis.qmd
Project Analysis Thursday PM

@@
 ```{r}
-p4 <- mydata %>% ggplot(aes(x=as.factor(Ownership.Type), y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
+
+mydata %>% count(Ownership.Type)
+
+mydata$owner <- substr(mydata$Ownership.Type, 1, 10)
+
+aggregate(mydata$Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, list(mydata$owner), FUN=mean, na.rm=TRUE)
+aggregate(mydata$HAI_039_RS_Rate, list(mydata$owner), FUN=mean)
+
+p4 <- mydata %>% ggplot(aes(x=as.factor(owner), y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
 plot(p4)
-figure_file = here("results","staff_hai_ownership.png")
+figure_file = here("results","hai_ownership.png")
 ggsave(filename = figure_file, plot=p4)
+
+view(mydata)
+p5 <- mydata %>% ggplot(aes(x=as.factor(owner), y=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_boxplot() + geom_smooth(method='lm')
+plot(p5)
+figure_file = here("results","staff_ownership.png")
+ggsave(filename = figure_file, plot=p5)
+
 ```
@@
 
-#Applying the model to unseen test data
-
-predict(model_fit2, test_data)
-
-model_aug2\<- augment(model_fit1, test_data)
-
-#pull RS HAI Rate and predictions
-
-temps \<- model_aug2 %\>% pull(HAI_039_RS_Rate)
-
-pred \<- model_aug2 %\>% pull(.pred)
-
-#run rmse
-
-rmse_vec(temps, pred)
-
 ```{r}
@@
 
-recipe_1 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)
+recipe_1 <- recipe(HAI_039_RS_Rate ~ owner + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)
 
-recipe_2 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + as.factor(Staffing.Rating), data= train_data)
+recipe_2 <- recipe(HAI_039_RS_Rate ~ owner + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + Staffing.Rating, data= train_data)
 
-recipe_3 <- recipe(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + as.factor(Staffing.Rating), data= train_data)
+recipe_3 <- recipe(HAI_039_RS_Rate ~ owner + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score + Staffing.Rating, data= train_data)
 

### 2023-05-05 products/manuscript/Manuscript.qmd
Project Analysis Thursday PM

@@
 
+## Hypothesis
+
+The hypothesis of this study is that nursing homes with higher staffing rates and higher five-star staffing rate will be associated with lower healthcare associated infections requiring hospitalizations.
+We further hypothesize that models will show that actual staffing rates will be a better variable for predicting infection rates than five-star staffing ratings.
+
 {{< pagebreak >}}
@@
 The HAI requiring hospitalization metric is also given as a categorical variable, where the nursing home is compared to all other nursing homes in that time period and categorized as average, significantly worse than average, or significantly better than average.
+
+## Exploratory analysis
+
+First, we looked through the distribution of all the variables we may utilize in our study.
+Visually, we examined histograms, box plots, and other representations of the distribution.
+We then looked at how our continuous variables were distributed by the categorical staffing star metric, or by the HAI rate categorization.
+
+In our exploratory analysis, we noticed missing values for some variables. Otherwise values were plausible and in range.
 
@@
 
-From all data, we identified 10,502 nursing homes with complete data to include in our analysis. Complete description of the cohort included in the study in the Supplementary Files.
+Our exploratory analysis showed us fairly normal distributions of the risk standardized HAI requiring hospitalization rate and staffing variables.
+We noticed expected and appropriate correlation with the continuous rate variables and their categorical rankings.
+For example, as the number of staffing hours per resident per day increased, the staffing star rating also increased.
 
-Distributions of all variables are included in the Appendix.
+For interpretation, we reduced the \>10 ownership categories into Non-profit, For-profit, and Government.
+We saw potentially lower HAI rates in non-profit (7.00) compared to for-profit (7.9) nursing homes, with government-owned homes falling in the middle (7.4).
+Conversely, staffing rates were highest in non-profit nursing homes (4.2) compared to for-profit nursing homes (3.5), again with government in the middle (3.8).
+
+From all data, we identified 10,502 nursing homes with complete data to include in our analysis. Some of these nursing homes were missing some staffing data.
 
@@
 
-We compared the three models using RMSE to determine which model best fit our data in the training dataset. The comparable RMSE rates are shown in Table 2.
-
-***#Add R Code here to show Table 2 (RMSE for different model choices 1,2,3)***
-
+In our exploratory analysis, we compared the three models using RMSE to determine which model best fit our data in the training dataset.
 We selected model 1, containing only the staffing rating as our best fitting model.When we applied model 1 to our test data we found that the predicted and observed values were fairly similar (Figure 5) and the RMSE was 1.79.
 
-***#Add R Code here to show Figure 5***
+```{r}
+#| label: fig-result5
+#| fig-cap: "Scatter plot of observed versus predicted values in test data"
+#| echo: FALSE
+knitr::include_graphics(here("results","scatter_model.png"))
+```
+
+###
+
+The parameter estimates from the model applied to the full dataset are shown in Table 2.
+
+```{r}
+#| label: Table 2
+#| fig-cap: "Table showing parameter estimates from selected model on full dataset"
+#| echo: FALSE
+resulttable2=readRDS("../../results/table2.RDS")
+knitr::kable(resulttable2)
+```
 
@@
 This gives us confidence that our conclusions from model 1 are robust.
-
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable)
-```
 
@@
 
-*What are the main take-home messages?*
-
-*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
+This analysis provides evidence that the risk standardized HAI metric may be able to elucidate important differences in quality of care in a nursing home setting.
+Further, it suggests that the five star staffing rating may be more informative than the staffing rate alone at predicting nursing homes with higher HAI rates.
+Nursing homes with inadequate staff should be provided additional resources to help ensure quality staffing is available to help keep residents safer from unnecessary harm.
 
